MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?

### 2020-08-25 Tex/MSc Data Science Project.tex
Update graph caption and report updates

\documentclass[12pt]{article}
\usepackage{arydshln}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{mdframed}
\usepackage[margin=3cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{textcomp}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage[colorlinks=true]{hyperref}
\usepackage{cleveref}
\crefdefaultlabelformat{[#2#1#3]}
\begin{document}
\renewcommand\citeform[1]{[#1]}
%
\title{MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?}
\author{\\Ryan Hill MSci\\\\
Supervisor: Dr Hubie Chen\\\\
Birkbeck, University Of London\\\\
\texttt{rhill06@mail.bbk.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
\graphicspath{{images/}}
\begin{abstract}
	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in and attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image. We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work. Using an Inception-ResNet-v2 we achieve Top-1 accuracy of 24.7\% and a Top-3 accuracy of 45.4\% which is an improvement over previous results.
\end{abstract}
\begin{figure}[!b]
	\centering
	\includegraphics[scale=0.4]{bbk_logo.jpg}
\end{figure}

\clearpage
%
{\hypersetup{linkcolor=black}
\tableofcontents}
\thispagestyle{empty}
\clearpage
%
\setcounter{page}{1}
\section{Introduction} 
\label{sec:intro}
Convolutional neural networks (CNNs) have been used for image classification for the last 30 years\cite{Cun1989}, with ever more accurate, and more complex, model architectures being produced and tuned on benchmark datasets. CNNs have been used to identify clothing types, objects, music, and even identify artists of paintings, but as far as we are aware only once has someone before asked the question of the ability of a CNN to judge a book by its cover i.e. predict the genre of the book based on the cover image. We will discuss the work of this person in \cref{sub:Related Works}.

@@

In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image. The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Technology} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs. In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images. \Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising out work in \cref{sec:Conclusion}. Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.

\subsection{Related Works} 
\label{sub:Related Works} 
So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work completed by Iwana et al.\cite{KenjiIwana}. In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}. These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}. In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.

@@

Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly epochs than the original paper.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
Whilst in this work we do not detail the full design and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks. In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}. We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks. For this section and the remainder of this work, unless otherwise specified, we observe the following notational conventions:
\begin{itemize}
	\item Unless otherwise stated all tensors (including vectors and matrices) are indexed from 1.
	\item Roman indices will be used to reference specific components of a tensor within an equation, Greek indices will be used for dimensions to be operated over only.
	\item Dimensions of each tensor when referring to an image or feature map will be referred to in the order of height, width, (colour) channel (interchangeable with depth), including in the indices.
	\item Bold items will refer to full vectors of matrices, with it clear which one from the context.
\end{itemize}
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
\subsubsection{Activation functions}
Most layers of a neural network have some kind of activation function; these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range). Some normalisation is also usually applied which is discussed later. For our purposes there are only 3 activation functions that are relevant for this work.

@@

\subsubsection{Convolutional layers}
A convolutional layer, the key part of a neural network, is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN. A convolutional layer is defined by 3 key parameters, and the additional choice of padding. The 3 parameters that define a basic convolutional layer are the number of filters (analogous to the number of neurons in a dense layer) which is also the number of channels the output of that convolution will have, the size of the filters, and the stride (or step size) of the filter.

A filter (also called a kernel) is applied atop of the input, starting at the top left of the input, and pointwise multiplication is applied to each set of elements, then their sum is computed. In the case of multiple input channels the same filter is applied over all channel and their outputs are summed. This value is the the output value of the top left cell of the output tensor. The filter is then moved left by the number of elements defined by the stride; when the end of the row has been reached the filter is move down by the same stride. This process is repeated until the filter has been applied to the entire input. Once all filters have been applied this set of outputs is referred to as the feature map. This process is demonstrated in \cref{fig:conv_process} and quantified mathematically as

@@

\subsubsection{Pooling}
Pooling is an operation used within convolutional neural networks to reduce the size of channels/feature maps within the network. Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area. These operations tend to either be maximum or average and is determined either by the user of the CNN architecture. These are applied per-channel and no aggregation across channels is completed afterwards. The dimensions of the output after pooling can be calculated using the same method as before displayed in \cref{eq:feat_map2}. Traditionally (and in our work) pooling is completed with no padding. Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned.

@@

In transfer learning it is possible to take one of two approaches; the first is to use the existing model weight and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e. all but the fully connected layers at the end, and just train the classification top of the model. In our work we took the second approach as given the size of the models we are using training the full model would have been computationally expensive, and given the relatively small amount of training data would also have when compared to the number of parameters in the model it would have been unlikely to have improved accuracy.
% subsection Transfer_Learning end

\subsection{Activation Maximisation} 
\label{sub:Activation_Maximisation} 
TBC: Describe the theory and/or maths of activation maximisation and the methods used within the package we use the regularisations applied. 
% subsection Activation_Maximisation end 
% section Theory end 
\section{Data collection and pre-processing} 
\label{sec:Data_collection_and_pre-processing} 
\subsection{Image Download and Manual Review} 
\label{sub:Image Download and Manual Review} 
The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}. The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway. Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e. the genre classification.

Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked. At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, an example of such an image can be seen in \cref{fig:bad_cover}. The terms used in a regex search of the titles were \emph{``boxed set'', ``boxset'', ``box set'', ``anthology'', ``bundle'', ``\#-book'', and ``\# book''}. 264 records within the bookcover30 dataset matched at least one of these terms and were extracted into a new folder for manual review. The process was somewhat subjective but as a rule of thumb we rejected any image that was not a single front facing cover, that contained pictures of multiple books and was not an arrangement of covers; single covers that seemed to be specifically designed for a boxset were considered acceptable. In total 94 records needed to be removed due to these rules. To preserve the class balance of the train and test datasets, new records were chosen from the book32 dataset by making like-for-like swaps of the same class; any new covers were also reviewed to ensure they were not boxsets. No verification of the class assignment to images was performed so any errors that existed in the original data would have persisted and we continue to use the single class chosen by the original authors in the case of multi-class books.

@@

We chose to split the training data out into training and validation data using a 90/10 random split per class to ensure class balance in all of the training, validation, and test sets. In total the training set compromised of 46,170 records, the validation set 5130 records, and the test set 5700 records; with 1,539, 171, and 190 records per class respectively.

@@

We next used 3 different image preparation methods on the datasets to attempt to identify the method that would lead to the best accuracy within our models. At this stage all methods output an image 299x299 pixels as this was the largest size we would need for any future models. By reducing the image size twice, first now and then again when using a specific model, we were potentially losing some information that would lead to a less accurate model; however we chose to do this to speed up training and testing by having to load much smaller data in compared to the original full sized images. Before pre-processing was done we collected the initial shape and average RGB values of each image for further analysis later.

@@

% section Image_Pre-processing_Technique end 
% section Data_Collection_and_pre-processing end 
\section{Model Training} 
\label{sec:Model_Training} 
\subsection{Configuration} 
\label{sub:Configuration} 
All model architectures consisted of their respective base CNN structure as detailed in \cref{sec:CNN_Architectures} followed by a global average pooling 2D layer to reduce the dimensionality of our tensor and to ensure a flat 1D vector which is fully connected to a 30 neuron output layer with a softmax activation. All models were also trained using the same setup in terms of hyperparameters, optimisers and loss functions; the loss function was sparse categorical cross entropy (sparse due to the usage of a non one-hot encoded target vector), the optimiser was the Adam optimiser \cite{Kingma2015}. The optimiser was used with default hyperparameters as defined within the TensorFlow implementation, the most important being the learning rate with was left at the initial value of 0.001. The reason for the usage of the default hyperparameters and not using a more advanced custom optimiser, such as WAME \cite{Mosca2017}, was due to the constraints on access time to computing power. As is the training already had to be completed across multiple sessions and doing any tuning would have resulted in too large a time required to produce meaningful results; Adam produces decent results with the defaults and so they were used in this work.

The models themselves, MobileNetV2, Inception-ResNet-v2, and ResNeXt50 were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}. MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResNet-v2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures. These models, combined with the original work done by Iwana et al, span a range of complexity and training speed allowing for a good comparison between performance.

The models were training once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point. No cross-validation was done due to resource limitations.
% subsection Configuration end 
\subsection{Training Performance} 
\label{sub:Training_Performance} 
The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most.

The performance of the models throughout training can be seen in \cref{fig:train_perf} where whilst training accuracy continues to increase over time, the validation accuracy doesn't see as much consistent movement which suggests the model is starting to overfit to the training data as opposed to improving the identification of intrinsic underlying features within the images themselves. Iwana does not report their training accuracy so we are not able to compare the performance of our training with theirs.

@@

Iwana et al. trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs. It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.

MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNet-v2 continued to see improvement until just before the 50th epoch, ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
% subsection Training_Performance end

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison. A visual representation of this for our models only is shown in \cref{fig:test_perf}. As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset did not contain the set of records; but for a comparison it is suitable. Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account. MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres. ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs. This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.

@@

@@

	\caption{Top-1 Per class and total precision for each model}

@@

First, as displayed in \cref{fig:col_distributions}, we can see that whilst there is a wide range of average RGB values and width/aspect ratio (height is almost entirely consistent across genres), there are no standout values in colour alone that allow for easy genre classification. Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some of the more wider books than many other categories. Whilst none of this alone would be enough to provide good classification, it does suggest that explicit inclusion of dimensional information in a model could improve the performance when combined with the existing output.

@@

The average overall colour per genre is presented in \cref{fig:avg_col} which does suggest some slight differences between the genres; similar to the callouts made by Iwana, Sci-Fi and Fantasy seem to be darker when compared to something like Cookbooks, Food and Wine. We can compare the average colour of the true classes vs those predicted by each model and visualise how these correlate with the recall per class as shown in \cref{fig:col_distance}. Here the colour distance is defined as

\begin{equation}
	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
\end{equation}
where R/G/B are the average red, green, and blue values for each predicted and class respectively. We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class. This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements. This supports the claims made by Iwana in their work.

@@

\subsection{Discussion \& Future Work} 
\label{sub:Discussion} 
Overall the Inception-ResNet-v2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results. The counterpoint to this is that the AlexNet model as trained by Iwana et al. was trained for 450,000 epochs where out model was trained for less than 50 epochs total. We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greatly difference between the results. Despite training on slightly less data (due to the validation set), we can see that more modern and powerful CNNs are able to predict book genres with slightly improved accuracy compared to older methods.

>>>>

\documentclass[12pt]{article}
\usepackage{arydshln}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{mdframed}
\usepackage[margin=2cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{textcomp}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage[colorlinks=true]{hyperref}
\usepackage{cleveref}
\crefdefaultlabelformat{[#2#1#3]}
\begin{document}
\renewcommand\citeform[1]{[#1]}
%
\title{MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?}
\author{\\Ryan Hill MSci\\\\
Supervisor: Dr Hubie Chen\\\\
Birkbeck, University Of London\\\\
\texttt{rhill06@mail.bbk.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
\graphicspath{{images/}}
\begin{abstract}
	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image. We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work. Using an Inception-ResNet-v2 we achieve Top-1 accuracy of 24.7\% and a Top-3 accuracy of 45.4\% which is an improvement over previous results.
\end{abstract}
\begin{figure}[!b]
	\centering
	\includegraphics[scale=0.4]{bbk_logo.jpg}
\end{figure}

\clearpage
%
{\hypersetup{linkcolor=black}
\tableofcontents}
\thispagestyle{empty}
\clearpage
%
\setcounter{page}{1}
\section{Introduction} 
\label{sec:intro}
Convolutional neural networks (CNNs) have been used for image classification for the last 30 years\cite{Cun1989}, with ever more accurate, and more complex, model architectures being produced and tuned on benchmark datasets. CNNs have been used to identify clothing types, objects, music, and even identify artists of paintings; but as far as we are aware only once has someone tested asked the question of the ability of a CNN to judge a book by its cover i.e. predict the genre of the book based on the cover image. We will discuss the work of this person in \cref{sub:Related Works}.

@@

In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image. The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs. In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images. \Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising out work in \cref{sec:Conclusion}. Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.

\subsection{Related Works} 
\label{sub:Related Works} 
So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work completed by Iwana et al.\cite{KenjiIwana}. In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}. These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}. In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% Top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.

@@

Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly epochs than the original paper.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
Whilst in this work we do not detail the full design (beyond that provided in \cref{sec:CNN_Architectures}) and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks. In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}. We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks.
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
\subsubsection{Activation functions}
Most layers of a neural network have some kind of activation function; these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range). Some normalisation is also usually applied which is discussed later. For our purposes there are only 3 activation functions that are relevant for this work.

@@

\subsubsection{Convolutional layers}
A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN. A convolutional layer is defined by 3 key parameters, and the additional choice of padding. The 3 parameters that define a basic convolutional layer are the number of filters (analogous to the number of neurons in a dense layer) which is also the number of channels the output of that convolution will have, the size of the filters, and the stride (or step size) of the filter.

A filter (also called a kernel) is applied atop of the input, starting at the top left of the input, and pointwise multiplication is applied to each set of elements, then their sum is computed. In the case of multiple input channels the same filter is applied over all channel and their outputs are summed. This value is the the output value of the top left cell of the output tensor. The filter is then moved left by the number of elements defined by the stride; when the end of the row has been reached the filter is moved down by the same stride. This process is repeated until the filter has been applied to the entire input. Once all filters have been applied this set of outputs is referred to as the feature map. This process is demonstrated in \cref{fig:conv_process} and quantified mathematically as

@@

\subsubsection{Pooling}
Pooling is an operation used within convolutional neural networks to reduce the size of channels/feature maps within the network. Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area. These operations tend to either be maximum or average and are determined either by the user or the CNN architecture. These are applied per-channel and no aggregation across channels is completed afterwards. The dimensions of the output after pooling can be calculated using the same method as before displayed in \cref{eq:feat_map2}. Traditionally (and in our work) pooling is completed with no padding. Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned.

@@

In transfer learning it is possible to take one of two approaches; the first is to use the existing model weight and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e. all but the fully connected layers at the end, and just train the classification top of the model. In our work we took the second approach as, given the size of the models we are using, training the full model would have been computationally expensive, and given the relatively small amount of training data would also have, when compared to the number of parameters in the model, been unlikely to have improved accuracy.
% subsection Transfer_Learning end

\subsection{Activation Maximisation} 
\label{sub:Activation_Maximisation} 
Activation maximisation is a technique that can be applied to neural networks, in particular CNNs, to help understand what an idealised input for a given output the network would expect. Is is often combined with saliency or attention maps to identify the area(s) of a specific image that the CNN is drawing on for the classification. Activation maximisation is used instead to produce an input for the network that maximises the activation of the target class.

Given the potentially complex networks that can be reviewed using this technique the theory itself remains quite simple in nature. Given a trained model, as defined by the architecture and parameters, that ends with a fully connected dense layer the first step is to freeze all the parameters for the model i.e. the weights and biases. Next, the activation function in the final layer (most likely a softmax function) should be replaced with a linear function instead. The reason for this replacement is subtle but important; due to the softmax function taking as an input the entire vector of values it is possible to maximise a single value by minimising the other values. If we were to continue to use the softmax it could be the case that in trying to generate an image that most represents class A, we instead generate an image the least represents all other classes; by using a linear activation instead we avoid this issue entirely.

Finally, a random input, $\bm{x}$ is chosen at the start and is evolved using backpropagation using the learning rule defined as

\begin{equation}
	\label{eq:act_max}
	\Delta x_i = -\eta\frac{\partial \hat{y}_t}{\partial x_i} + \mathcal{R}
\end{equation}
where $\eta$ is the learning rate which in this work was set at 1, $\mathcal{R}$ is any regularisation terms, and $\hat{y}_t$ is the activation value of the target class. Simply put, change the values for the input in a way that increases the activation for that class within some constraints. One regularisation term used in this work is L2norm. It is defined as the root mean squared value of the vector and in this case helps ensure to generate a smooth image, penalising edges.
% subsection Activation_Maximisation end 
% section Theory end 
\section{Data collection and pre-processing} 
\label{sec:Data_collection_and_pre-processing} 
\subsection{Image Download and Manual Review} 
\label{sub:Image Download and Manual Review} 
The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}. The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway. Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e. the genre classification. Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.

At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, an example of such an image can be seen in \cref{fig:bad_cover}. The terms used in a regex search of the titles were \emph{``boxed set'', ``boxset'', ``box set'', ``anthology'', ``bundle'', ``\#-book'', and ``\# book''}. 264 records within the bookcover30 dataset matched at least one of these terms and were extracted into a new folder for manual review. The process was somewhat subjective but as a rule of thumb we rejected any image that was not a single front facing cover or that contained pictures of multiple books and was not an arrangement of covers; single covers that seemed to be specifically designed for a boxset were considered acceptable. In total 94 records needed to be removed due to these rules. To preserve the class balance of the train and test datasets, new records were chosen from the book32 dataset by making like-for-like swaps of the same class; any new covers were also reviewed to ensure they were not boxsets. No verification of the class assignment to images was performed so any errors that existed in the original data would have persisted and we continue to use the single class chosen by the original authors in the case of multi-class books.

@@

We chose to split the training data out into training and validation data using a 90/10 random split per class to ensure class balance in all of the training, validation, and test sets. In total the training set compromised of 46,170 records, the validation set 5,130 records, and the test set 5,700 records; with 1,539, 171, and 190 records per class respectively.

@@

We next used 3 different image preparation methods on the datasets to attempt to identify the method that would lead to the best accuracy within our models. At this stage all methods output an image of 299x299 pixels as this was the largest size we would need for any future models. By reducing the image size twice, first now and then again when using a specific model, we were potentially losing some information that would lead to a less accurate model; however we chose to do this to speed up training and testing by having to load much smaller data in compared to the original full sized images. Before pre-processing was done we collected the initial shape and average RGB values of each image for further analysis later in \cref{sub:Further_Analysis}.

@@

% section Image_Pre-processing_Technique end 
% section Data_Collection_and_pre-processing end 
\section{Model Training} 
\label{sec:Model_Training} 
\subsection{Configuration} 
\label{sub:Configuration} 
All model architectures consisted of their respective base CNN structure as detailed in \cref{sec:CNN_Architectures} followed by a global average pooling 2D layer to reduce the dimensionality of our tensor and to ensure a flat 1D vector, this is then fully connected to a 30 neuron output layer with a softmax activation. All models were also trained using the same setup in terms of hyperparameters, optimisers and loss functions; the loss function was sparse categorical cross entropy (sparse due to the usage of a non one-hot encoded target vector), the optimiser was the Adam optimiser \cite{Kingma2015}. The optimiser was used with default hyperparameters as defined within the TensorFlow implementation, the most important being the learning rate with was left at the initial value of 0.001. The reason for the usage of the default hyperparameters and not using a more advanced custom optimiser, such as WAME \cite{Mosca2017}, was due to the constraints on access time to computing power. As is the training already had to be completed across multiple sessions and doing any tuning would have resulted in too large a time required to produce meaningful results; Adam produces decent results with the defaults and so they were used in this work.

The models themselves, MobileNetV2 \cite{Sandler2018}, Inception-ResNet-v2 \cite{Szegedy2016}, and ResNeXt50 \cite{Xie2016} were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}. MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResNet-v2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures. These models, combined with the original work done by Iwana et al, span a range of complexity and training speed allowing for a good comparison between performance.

The models were training once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point. No cross-validation was done due to resource limitations.
% subsection Configuration end 
\subsection{Training Performance} 
\label{sub:Training_Performance} 
The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most. The training was completed over multiple sessions due to timeout constraints of the Google Colab platform and the speed of training even using a GPU instance; during each session whenever the validation accuracy of an epoch outperformed the previous best this model was saved with the epoch identifier to Google Drive. When a new session was started the weights and biases of this snapshot were loaded directly in and training we restarted from the same epoch. This does mean that there may be slightly differences compared to had the model just been trained in a single session, but these effects should be minimal.

The performance of the models throughout training can be seen in \cref{fig:train_perf} where, whilst training accuracy continues to increase over time, the validation accuracy doesn't see as much consistent movement which suggests the model is starting to overfit to the training data as opposed to improving the identification of intrinsic underlying features within the images themselves. Iwana does not report their training accuracy so we are not able to compare the performance of our training with theirs.

@@

Iwana et al. trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, order of magnitudes more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs. It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.

MobileNetV2 achieved the best validation accuracy early on in its training, where Inception-ResNet-v2 continued to see improvement until just before the 50th epoch, ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance; the accuracy of ResNeXt50 at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
% subsection Training_Performance end

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison. A visual representation of this for our models only is shown in \cref{fig:test_perf}. As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the fill set of records; but for a comparison it is suitable. Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account. MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres. ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs. This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.

@@

	%\resizebox{\textwidth}{!}{%

@@

	%}
	\caption{Top-1 Per class, and total, precision for each model \\ \tiny{Missing values represent no predictions made in that class over the test set}}

@@

First, as displayed in \cref{fig:col_distributions}, we can see that whilst there is a wide range of average RGB values and width/aspect ratio (height is almost entirely consistent across genres), there are no standout values in colour alone that allow for easy genre classification. Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some wider books when compared to other categories. Whilst none of this alone would be enough to provide good classification, it does suggest that explicit inclusion of dimensional information in a model could improve the performance when combined with the existing output.

@@

The average overall colour per genre is presented in \cref{fig:avg_col} which does suggest some slight differences between the genres; similar to the callouts made by Iwana, Sci-Fi and Fantasy seem to be darker when compared to something like Cookbooks, Food and Wine. We can compare the average colour of the true classes against those predicted by each model and visualise how these correlate with the accuracy per class as shown in \cref{fig:col_distance}. Here the colour distance is defined as

\begin{equation}
	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
\end{equation}
where R/G/B are the average red, green, and blue values for each predicted and true class respectively. We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class. This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements. This supports the claims made by Iwana in their work.

@@

\subsection{Discussion \& Future Work} 
\label{sub:Discussion} 
Overall the Inception-ResNet-v2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results. The counterpoint to this is that the AlexNet model as trained by Iwana et al. was trained for 450,000 epochs where our model was trained for less than 50 epochs total. We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greater difference between the results. Despite training on slightly less data (due to the validation set), we can see that more modern and powerful CNNs are able to predict book genres with slightly improved accuracy compared to older methods.
